Are mental properties supervenient on brain properties?

### 2010-11-17 supervenience.tex
tried to make readable for others

%\input{/Users/joshyv/Research/misc/latex_paper.tex}
\documentclass{article}
\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{color}
% \usepackage{ulem}
% \usepackage{color}

% \newcommand{\loo}{$L^{(1)}_{h; \mD_n}$}
% \newcommand{\conv}{\rightarrow}
% \newcommand{\Real}{\mathbb{R}}
% \providecommand{\tr}[1]{\textcolor{red}{#1}}

\newcommand{\mB}{\mathcal{B}}
\newcommand{\mD}{\mathcal{D}}
\newcommand{\mM}{\mathcal{M}}
% \newcommand{\mN}{\mathcal{N}}

\newcommand{\hL}{\widehat{L}}
\newcommand{\MeB}{\mM \overset{\varepsilon}{{\sim}}_F \mB}
\newcommand{\MnoteB}{\mM \overset{\varepsilon}{{\not\sim}}_F \mB}
\providecommand{\tr}[1]{\textcolor{black}{#1}}
\providecommand{\norm}[1]{\left \lVert#1 \right  \rVert}
\newcommand{\T}{^{\ensuremath{\mathsf{T}}}}           % transpose

\newtheorem{defi}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{thex}{Thought experiment}
\lhead{Vogelstein JT, et al}
\rhead{Neurocognitive Graph Theory}

@@

\author{Joshua T. Vogelstein$^1$, R. Jacob Vogelstein$^{1,2}$, Carey E. Priebe$^1$\\
$^1$Department of Applied Mathematics and Sciences, \\ Johns Hopkins University, Baltimore, MD, 21218,\\ $^2$National Security Technology Department, \\ Johns Hopkins University Applied Physics Laboratory, Laurel, MD 20723}

@@

To address the question of whether the mind supervenes on the brain through empirical means, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to scientific methods and statistical analysis. To elucidate this approach, we posit a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of $\varepsilon$-supervenience. The physical property of the brain employed in this analysis is the graph describing brain connectivity (i.e., the \emph{connectome}), and $\varepsilon$-supervenience allows us to determine whether a particular mental property can be inferred from one's connectome to within any given misclassification rate $\varepsilon > 0$, regardless of the relationship between the two. In addition to the theoretical results, we show via simulation that given reasonable assumptions about class conditional probabilities and the amount of data available, the thought experiment can actually be conducted on a simple organism, \emph{Caenorhabditis elegans}, with currently available technology.

The potential significance of this work can be divided into distinct disciplines.  To the philosopher, this work demonstrates that philosophical conjectures can be morphed into statistical hypotheses, amenable to experimental investigations, allowing the philosopher to add empirical support to their rational arguments.  To the statistician, herein lies the first proof to our knowledge of the existence of a universally consistent classifier on graphs, and a constructivist one at that.  To the neuroscientist, a theoretically possible experiment is proposed to garnish support for a hypothesis that is widely believed: that mental properties supervene on brain properties.

@@

@@

\tr{The primary contributions of this work are as follows.  First, a notion of supervenience amenable to empirical investigation is formally introduced.  This renders the mind-brain dualism debate a hypothesis, rather than an assumption.  Second, in addition to expanding the space of questions amenable to hypothesis testing, we also demonstrate the the limits of hypothesis testing.  Third we posit a very general model of brains and their associated mental properties that admits statistical analysis in a graph theoretical and statistical framework.  Fourth, we prove that this formulation admits a universally consistent classifier that is guaranteed to find the relationship between minds and brains, if one exists.  Fifth we demonstrate through simulation that the proposed universally consistent classifier has reasonable convergence properties on simulated brain-graph data.}

Let $\mB$ be the observation space for some physical property, such as brain connectivity structure (i.e., connectome; see \cite{SpornsKotter05,LichtmanSanes08,Seung09}). Let $\mM$ be the (finite) indicator space for some mental property, such as knowing calculus. Thus, for $b \in \mB$ and $m \in \mM$, the pair $(b,m)$ represents a brain property/mind property pair.

%Let $\{(B_i,M_i)\}_{i=1}^n \overset{iid}{\sim} F_{BM}$
Let $(B,M), (B_1,M_1),\cdots,(B_n,M_n)$ be random observation pairs taking their values in $\mB \times \mM$, independently and identically distributed according to some joint probability distribution $F=F_{BM}$. Abusing notation to conceptually identify the properties with their spaces, the statistical supervenience relation $\mM \overset{S}{\sim}_F \mB$ says that $M_i \neq M_j \implies B_i \neq B_j$ (almost surely; where $\implies$ does not suggest causation). That is, observing $B=b$ can allow us to assign $m$ to $M$. While previously proposed notions of mind-brain-supervenience claim that all mental properties supervene on physical properties \cite{Kim95}, here we consider empirically investigating only whether a particular mental property $\mM$ statistically supervenes on a particular physical property $\mB$.

Let $g:\mB \rightarrow \mM$ be a classifier, which takes as input an observed brain connectivity structure $b$ and produces a classification $\widehat{m}=g(b)$ for the unobserved mental property $m$. The Bayes optimal classifier $g^*$ minimizes $L_{F}(g)$ over all classifiers, where $L_{F}(g) = P_{F}[g(B) \neq M]$ denotes the probability of misclassification for classifier $g$ under joint distribution $F=F_{BM}$. We can therefore rigorously define \textit{statistical supervenience}:

\begin{defi}
\label{def1} 
$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{BM}$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $L_{F}(g^*) = 0$.
\end{defi}

\tr{(Note that this definition does not imply a one-to-one mapping.)}  \tr{To allow for the possibility of only \emph{partial} supervenience,}  we relax the above statistical supervenience to define $\varepsilon$-\textit{supervenience}:

\begin{defi}
\label{def2}
Given $\varepsilon > 0$, $\mM$ is said to $\varepsilon$-\textit{supervene} on $\mB$ for distribution $F=F_{BM}$, denoted $\MeB$, if and only if $L_{F}(g^*) < \varepsilon$.
\end{defi}

\tr{Unfortunately, in general $F$ is unknown, but can be estimated from the data. Therefore,} generalizing the concept of a classifier $g$ to allow consideration of training data, consider $g_n:\mB \times (\mB \times \mM)^n \rightarrow \mM$ which takes as input an observed brain connectivity structure $b$ and $n$ training pairs $\tr{\vec{d}_n=}(b_1,m_1),\cdots,(b_n,m_n)$ and produces a classification $\widehat{m}=g_n(b;\tr{\vec{d}_n})$. Let $L_{F}(g_n) = E[P_{F}[g_n(B;\tr{\vec{D}_n}) \neq M|\tr{\vec{D}_n}]]$.

Consider the problem of testing for $\varepsilon$-supervenience. Let the null hypothesis be given by $H_0: L_{F}(g_n) \geq \varepsilon$ so that if we reject at level $\alpha >0$ in favor of the alternative hypothesis $H_A: L_{F}(g_n) < \varepsilon$ then we can conclude, with $100(1-\alpha$)\% confidence, that $\MeB$. Letting $\hL^{n'}_{F}(g_n)$ denote the hold-out estimate of misclassification performance based on $n'$ test observations, we note that $\hL^{n'}_{F}(g_n)$ is distributed $Binomial(n',L_{F}(g_n))$. The test rejects for small $\hL^{n'}_{F}(g_n)$. The level $\alpha$ critical value $c_{\alpha}(n',\varepsilon)$ is available under the least favorable distribution $Binomial(n',\varepsilon)$. Furthermore, $\MeB$ implies $L_{F}(g^*) < \varepsilon$, and thus if $g_n$ is a {\it consistent} classifier for $F=F_{BM}$ --- that is, if $\lim_n L_{F}(g_n) = L_{F}(g^*)$ --- then the power of this test (the probability of rejecting when in fact the alternative is true) goes to unity as $n,n' \rightarrow \infty$. Thus we have an inference procedure: \begin{thm} \label{thm1} Given $\alpha > 0$, we can test $\MeB$ so that rejection implies $\MeB$ holds with probability greater than or equal to $1-\alpha$. Furthermore, given a consistent classifier the power of the test converges to unity. \end{thm}

Since the joint distribution $F=F_{BM}$ is unknown, the utility of Theorem \ref{thm1} requires that $g_n$ be a {\it universally consistent} classifier --- that is, $\lim_n L_{F}(g_n) = L_{F}(g^*)$ for all distributions $F=F_{BM}$. Unfortunately, the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ depends on the (unknown) distribution $F=F_{BM}$ \cite{DGL96}. Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}. For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation. Thus, without restrictions on $F_{BM}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}. Given these caveats, consider the following thought experiment:

\begin{thex}
Let the physical property under consideration be brain connectivity structure (``connectome''), so $b$ is a graph \tr{(or, network)} with vertices representing neurons (or neuroanatomical regions) and edges representing connections between neurons (or white matter tracts). Further let $\mB$, the observation space, be the collection of all graphs on a finite number of vertices, and let $|\mB|$ be countable. Now, imagine collecting very large amounts of very accurate independent and identically distributed brain-graph data and the associated mental property indicators. A $k_n$-nearest neighbor classifier using an isomorphism-matching Frobenius norm is universally consistent (see Appendix 1 for proof). Therefore, Theorem \ref{thm1} applies and the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mental/brain property pair, if indeed $\varepsilon$-supervenience holds. This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes. Furthermore, Appendix 1 also extends the proof to deal with other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite.  
\end{thex}

While the above thought experiment addresses the question of $\varepsilon$-supervenience, it does not address causality. Assuming we have confirmed $\MeB$ for a particular mental/brain property pair with confidence level $\alpha$, then morphing the brain (by altering edges) could be used to determine whether the relation is in fact causal.

Practical issues regarding actually conducting the above thought experiment include: (1) as stated, we must consider the space $\mB$ to be the quotient space of graphs mod graph isomorphism, unless the vertices are {\it labeled}; (2) a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under our Frobenius norm (or other norm; see Appendix 1 for details) may have a rate of convergence so slow and a computational demand so high as to be impractical; and (3) collecting enough sufficiently accurate independent and identically distributed brain-graph data and the associated mental property indicators may be beyond current technological capabilities. Regardless, related experimental work includes collecting various types of brain graph data \cite{WhiteBrenner86, DenkHorstmann04, BriggmanDenk06} and various approaches to inference on brain graphs \cite{MackeBorst08, Mishchenko09, LuLichtman09}, suggesting feasibility of such an experiment in the near future (see Appendix 2 for a simulated example of a feasible experiment). Nevertheless, our thought experiment suggests that we can hope to determine that a given mental property under consideration $\varepsilon$-supervenes on a brain's connectivity structure. \tr{This thought experiment, together with (i) the formal definition of $\varepsilon$-supervenience, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first \emph{proof} (to our knowledge) that empirically investigating supervenience is at least theoretically possible.}

%% == end of paper:

%% Optional Materials and Methods Section
%% The Materials and Methods section header will be added automatically.

%% Enter any subheads and the Materials and Methods text below.
%\begin{materials}
% Materials text
%\end{materials}

%% Optional Appendix or Appendices
%% \appendix Appendix text...
%% or, for appendix with title, use square brackets:
%% \appendix[Appendix Title]

\section*{Appendix 1: $k_n$-nearest neighbor universal consistency for graphs}
\label{proof}

Assume first that all graphs are simple \tr{(meaning undirected with no loops and binary binary edges)}, on the same set of vertices, and that the graphs are labeled so that we know which vertex in one graph corresponds to which vertex in another. Then the Frobenius distance function $d(b_1,b_2)$ can be written in terms of the associated adjacency matrices $A_1$ and $A_2$: $d(b_1,b_2) = ||A_1-A_2||_F$. If the graphs are identical, then $d(b_1,b_2) = 0$, and if the graphs are different, then $d(b_1,b_2) \geq 1$. Since the space $\mB$ is finite, $n$ large enough guarantees that with probability approaching unity at least $k_n$ training samples coincide with each atom, so long as $k_n/n \rightarrow 0$. Then $k_n \rightarrow \infty$ guarantees that the nearest neighbor vote-winner for each atom will eventually coincide with Bayes' choice, yielding universal consistency.

In the foregoing argument, there exists a smallest non-zero atomic probability $p_{min}$, and ``$n$ large enough'' is driven by this probability. Generalizing to countable $\mB$ with discrete weights, we see that given $\delta > 0$, there is a finite set $S$ with $P[S]>1-\delta$ and smallest atomic probability $p_{min}$, so that $L_{F}(g_n) \rightarrow c \leq L_{F}(g^*) + \delta$, yielding universal consistency.

If the graphs may have different numbers of vertices, and are unlabeled, we consider the isomorphism-matching Frobenius norm. Assume without loss of generality that $b_1$ has at least as many vertices as $b_2$, and write $A_2^P$ for the adjacency matrix associated with $b_2$ ``padded'' to include extra isolated vertices so that $A_2^P$ is the same size as $A_1$. Then $d(b_1,b_2) = \min_Q ||Q A_1 Q^T - A_2^P||_F$ where the minimum is taken over all permutation matrices \cite{HornJohnson90}. Under the equivalence relation induced by this isomorphism-matching, the foregoing universal consistency argument holds.
%\tr{Trivially, this argument also applies to graphs with (self-) loops, and
%directed, colored, or weighted edges, and hypergraphs.}

Several points of note:  isolated vertices are ignored in our equivalence relation; the class-conditional signal is entirely encompassed by the connectivity structure; the graph isomorphism problem is computationally hard \cite{ConroyLouis97,ZaslavskiyVert08}; and the argument employed here does not capture the concept of ``nearness implies likelihood of similar class''---we simply rely on atomic behavior.

Finally, the above proof can be straightforwardly generalized to utilize any matrix norm, $\norm{b_1-b_2}_A=\norm{A(b_1-b_2)}$, assuming that $A A\T$ is positive definite (see \cite{DGL96} pg. 455 for proof when $b_i$'s are vectors).  Furthermore, we can relax the constraint that mental properties finite, that is, we can allow any $|\mM| = \infty$ (see \cite{Stone77} for proof when $b_i$'s are vectors).  The proofs for the case when $b$ is an adjacency matrix follows immediately upon first defining a bijective embedding of the matrix into a vector space, and then embedding each adjacency matrix in that space.

%\appendix
\section*{Appendix 2: Simulation} \label{simulation}

As an example of a feasible experiment, one may consider a species whose nervous system consists of the same (small) number of labeled neurons for each organism. {\it Caenorhabditis elegans} is believed to be such a species \cite{Durbin87}. The hermaphroditic C.~elegans' somatic nervous system consists of 279 interconnected neurons. While the graph with these neurons as vertices and edges defined by chemical synapses between neurons is not identical across individuals, it is reasonably consistent \cite{Durbin87}. Furthermore, these animals exhibit a rich behavioral repertoire that depends on circuit properties \cite{deBonoMaricq05}. Thus, one may design an experiment by describing the joint distribution $F_{BM}$ via class-conditional distributions $F_{B|M=m_j}$ for the C.~elegans brain-graph for two mental properties of interest, $m_0$ and $m_1$, along with the prior probability of class membership $P[M=m_1]$. Here the mental property corresponds to the C.~elegans exhibiting or not exhibiting a particular behavior (e.g., response to an odor).

>>>>

%\input{/Users/joshyv/Research/misc/latex_paper.tex}
\documentclass{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{cite}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{times}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{color}
\pagestyle{fancy}

\oddsidemargin=0.0in %%this makes the odd side margin go to the default of 1inch
\evensidemargin=0.0in
\textwidth=6.5in
\headwidth=6.5in
\textheight=9in %%sets the textwidth to 6.5, which leaves 1 for the remaining right margin with 8 1/2X11inch paper
\headheight=12pt
\topmargin=-0.25in
%\headheight=0in
%\headsep=0in
%\pagestyle{headings}

\usepackage{hyperref}
% \usepackage{ulem}
% \usepackage{color}

% \newcommand{\loo}{$L^{(1)}_{h; \mD_n}$}
\newcommand{\conv}{\rightarrow}
% \newcommand{\Real}{\mathbb{R}}
% \providecommand{\tr}[1]{\textcolor{red}{#1}}

\newcommand{\mB}{\mathcal{B}}
\newcommand{\mD}{\mathcal{D}}
\newcommand{\mM}{\mathcal{M}}
\newcommand{\PP}{\mathbb{P}}           % probability
\newcommand{\EE}{\mathbb{E}}           % expected value
\newcommand{\II}{\mathbb{I}}           % expected value
\newcommand{\Real}{\mathbb{R}}           % expected value

\newcommand{\del}{\delta}
\newcommand{\sig}{\sigma}
\newcommand{\lam}{\lambda}
\newcommand{\gam}{\gamma}
\newcommand{\eps}{\varepsilon}

\providecommand{\mc}[1]{\mathcal{#1}}
\providecommand{\mb}[1]{\boldsymbol{#1}}
\providecommand{\mbb}[1]{\mathbb{#1}}
\providecommand{\mv}[1]{\vec{#1}}
\providecommand{\mh}[1]{\widehat{#1}}
\providecommand{\mt}[1]{\widetilde{#1}}
\providecommand{\mhc}[1]{\hat{\mathcal{#1}}}
\providecommand{\mhb}[1]{\hat{\boldsymbol{#1}}}
\providecommand{\mvb}[1]{\vec{\boldsymbol{#1}}}
\providecommand{\mtb}[1]{\widetilde{\boldsymbol{#1}}}

\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}

% \newcommand{\mN}{\mathcal{N}}

\newcommand{\hL}{\widehat{L}}
\newcommand{\MeB}{\mM \overset{\varepsilon}{{\sim}}_{\PP} \mB}
\newcommand{\MsB}{\mM \overset{S}{\sim}_F \mB}
\newcommand{\MnoteB}{\mM \overset{\varepsilon}{{\not\sim}}_{\PP} \mB}
\providecommand{\tr}[1]{\textcolor{black}{#1}}
\providecommand{\norm}[1]{\left \lVert#1 \right  \rVert}
\newcommand{\T}{^{\ensuremath{\mathsf{T}}}}           % transpose

\newtheorem{defi}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{thex}{Thought experiment}
\lhead{Vogelstein JT, et al}
\rhead{Statistical Supervenience}

@@

\author{Joshua T. Vogelstein$^1$, R. Jacob Vogelstein$^2$, Carey E. Priebe$^1$\\
$^1$Department of Applied Mathematics \& Statistics, \\ Johns Hopkins University, Baltimore, MD, 21218,\\ $^2$National Security Technology Department, \\ Johns Hopkins University Applied Physics Laboratory, Laurel, MD 20723
}

@@

To enable addressing the question of whether the mind supervenes on the brain through empirical means, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to empirical investigations and statistical analysis. To elucidate this approach, we posit a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of $\varepsilon$-supervenience. The physical property of the brain employed in this analysis is the graph describing brain connectivity (i.e., the \emph{connectome}), and $\varepsilon$-supervenience allows us to determine whether a particular mental property can be inferred from one's connectome to within any given misclassification rate $\varepsilon > 0$, regardless of the relationship between the two. %In addition to the theoretical results, we show via simulation that given reasonable assumptions about class conditional probabilities and the amount of data available, the thought experiment can actually be conducted on a simple organism, \emph{Caenorhabditis elegans}, with currently available technology.

The potential significance of this work can be divided into distinct disciplines.  To the philosopher, this work demonstrates that philosophical conjectures can be morphed into statistical hypotheses, amenable to experimental investigations, allowing the philosopher to add empirical support to their rational arguments.  To the statistician, this work points out the limitations of hypothesis testing in a novel domain.   To the neuroscientist, cognitive scientist and psychologist, this work indicates that much of contemporary research can be considered supervenience investigations, further motivating brain-graph related research paradigms.

@@

\section{Introduction}

@@

The primary contributions of this work are as follows.  A notion of supervenience amenable to empirical investigation is formally introduced.  This renders the mind-brain dualism debate a hypothesis, rather than an assumption, both expanding the space of questions amenable to hypothesis testing, and placing limits on this space.  Because hypothesis tests (implicitly sometimes) depend on a model, a very general model of brains and their associated mental properties is proposed.  Fortunately, this formulation admits universally consistent classifiers, that is, classifiers guaranteed to find the relationship between minds and brains, if one exists.  Many previous investigations relating brains and mental properties can therefore be considered $\eps$-supervenience hypothesis tests.  This paradigm, therefore, generalizes previous approaches, embedding them in a rigorous statistical framework, suggesting avenues for future research.

 % we demonstrate through simulation that the proposed universally consistent classifier has reasonable convergence properties on simulated brain-graph data. Finally, we discuss both contemporary and future applications of this framework to neurobiological studies in humans.

\section{Preliminaries} % (fold)
\label{sec:preliminaries}

\subsection{Supervenience} % (fold)
\label{sub:supervenience}

% subsection supervenience (end)
The intention in this work is to develop greater insight regarding the relationship between minds and brains, using statistical methods, with particular interest in notions of supervenience.  Modern philosophy of science suggests that adequate scientific explanations must account for the objects of investigation and their relationships \cite{Craver07}. Therefore, here we define minds, brains, and supervenience.

Let $b$ correspond to an agent's brain, which is a particular element from the set of all possible brains, $\mB$.  The set of possible brains $\mB$ is completely unrestricted, meaning that $\mB$ could be an infinite set, with arbitrarily complexity.  In particular, $b$ might represent the position, momentum, and type of each subatomic particle residing within the skull some agent.  Thus, each different $b \in \mB$ could correspond to some difference in the position, momentum, or type of at least one subatomic particle composing the brain.

Similarly, let $m$ correspond to an agent's mind, which is a particular element from the set of all possible minds, $\mM$.  The set of possible minds $\mM$ is also unrestricted. In particular, $m$ might represent all an agent's thoughts, beliefs, and preferences. Thus, each different $m \in \mM$ could correspond to a difference in at least one thought, belief, or preference.

The mind-brain supervenience conjecture is a relation between the set of mental states and the set of brain states.  Donald Davidson canonized this conjecture in 1970 with the following quote: \cite{Davidson70}
\begin{quotation}
\noindent supervenience might be taken to mean that there cannot be two events alike in all physical respects but differing in some mental respect, or that an object cannot alter in some mental respect without altering in some physical respect.
\end{quotation}
This conjecture may be concisely and formally stated: $m \neq m' \implies b \neq b' \, \forall (m,b), (m',b') \in \mc{M} \times \mc{B}$, where $(m,b)$ is a mind-brain pair, and $(m',b')$ is a different mind-brain pair.  While mind-brain supervenience is a relatively strong claim, importantly, one can imagine far stronger relations.

For instance, it may be the case that minds supervene on brains, but one cannot form an \emph{injective} relation from brains to minds.  An injective relation is any relation that preserves distinctness.  Thus if minds are injective on brains, then $b\neq b' \implies m \neq m' \, \forall (m,b), (m',b') \in \mc{M} \times \mc{B}$ (note that the directionality of the implication has been switched relative to supervenience). For instance, it might be the case that a brain could change without the mind changing.  Consider the case that a single subatomic particle shifts its position by a Plank length, changing brain state from $b$ to $b'$.  It is possible that the mental state supervening on brain state $b$ remains $m$, even after $b$ changes to $b'$.  In such a scenario, the mind might still supervene on the brain, but the relation from brains to minds is not injective. This argument also shows that supervenience is not necessarily a \emph{symmetric} relation.  Minds supervening on brains does not imply that brains supervene on minds.

% In this case, it would be said that both $M$ supervenes on $B$ and $B$ supervenes on $M$. But it is not necessarily the case that if supervenience holds in one direction that it does or does not hold in the other direction.

Second, it may be the case that minds supervene on brains, but that brains do not cause minds.  For instance, consider an analogy where $M$ and $B$ correspond to two coins being flipped, each possibly landing on heads or tails.  Further assume that every time one lands on heads so does the other, and every time one lands on tails, so do the other. This implies that $M$ supervenes on $B$, but assumes nothing about whether $M$ causes $B$, or $B$ causes $M$, or some exogenous force causes both.

% Supervenience is also not \emph{causality}.  Consider the coin flip example.  Nowhere did the notion of causality even arise.  It could be the case that $M$ causes $B$, or $B$ causes $M$, or some exogenous force causes both.  Thus, supervenience does not imply causality.

Third, supervenience does not imply \emph{identity}.  Consider, for example, acceleration and velocity.  Clearly, acceleration supervenes on velocity, as acceleration cannot change without velocity changing (assuming one does not consider gravity as acceleration).  Similarly, velocity supervenes on position, as velocity cannot change without position changing.  Therefore, acceleration supervenes on position, by the transitive property of supervenience, but it is not the case that a change in acceleration is equal to a change in position.  Rather, position can change with constant velocity, meaning without acceleration changing.

What supervenience does imply, however, is the following.   Imagine finding two different minds.  If $\MeB$, then the brains subvening under those two minds must be different.  In other words, there cannot be two different minds, either of which could supervene on a single brain.  Figure \ref{fig:rel} shows several possible relations between the sets of minds and brains.

\begin{figure}[htbp]
	\centering
		\includegraphics[width=1\linewidth]{supervenience_relations.pdf}
	\caption{Possible relations between minds and brains.  (A) Minds supervene on brains, and it so happens that there is a bijective relation from brains to minds.  (B) Minds supervene on brains, and it so happens that there is a surjective (a.k.a., onto) relation from brains to minds. (C) Minds are \emph{not} supervenient on brains, because two different minds supervene on the same brain.}
	\label{fig:rel}
\end{figure}

All of the above relations are \emph{logical} relations, not probabilistic relations. To facilitate both statistical analysis and empirical investigation, we project this supervenience notion into a statistical setting.  To proceed, we first define a \emph{model}, $\mc{P}=\{\PP[M,B]\}$, the set of all possible joint distributions, $\PP[M,B]$, that could describe the probabilistic relationship between minds and brains.  Each distribution $\PP[M,B]$ uniquely specifies the probability of any element $(m,b)$ occurring from the space of all possible elements $\mc{M} \times \mc{B}$ (also called the \emph{sample space}).%  The true distribution could in theory be any possible joint distribution from the model, $\PP[M,B] \in \mc{P}$, meaning that we only assume that it conforms to the \emph{probability axioms}:
% \begin{enumerate}
% 	\item $0 \leq \PP[B=b,M=m] \leq 1$ for all $(m,b) \in \mM \times \mB$
% 	\item $\PP[\Omega]=1$ and $\PP[\emptyset]=0$, where $\Omega$ is the whole sample sample, that is, all elements, $(m,b) \in \mc{M} \times \mc{B}$
% 	\item any countable sequence of pairwise disjoint elements, $(m_1,b_1), (m_2,b_2), \ldots$ satisfies \\
% 	\mbox{$\PP[(m_1,b_1) \cup (m_2,b_2) \cup \ldots] = \sum_i \PP[(m_i,b_i)]$}.
% \end{enumerate}
Given a distribution, one can then calculate any marginal or conditional distributions.  For instance, $\PP[B]=\int_{m \in \mM} \PP[M,B]dm$, or $\PP[M|B]=\PP[M,B]/\PP[B]$.  Having specified a model, statistical supervenience can be defined as follows:
\begin{defi}
\label{def1} 
$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $\PP[m \neq m' | b=b']=0 \, \forall (m,b), (m',b') \in \mc{M} \times \mc{B}$. Alternately, $\MeB$ can be written as $\PP[m = m' | b = b']=1  \, \forall (m,b), (m',b') \in \mc{M} \times \mc{B}$.
\end{defi}
Statistical supervenience is therefore a probabilistic relation on sets.  Note that statistical supervenience is distinct from statistical correlation.  \emph{Statistical correlation} between brain states and mental states is defined as $\rho_{MB}=\EE[(B-\mu_B)(M-\mu_M)]/(\sig_B \sig_M)$, where $\mu_X$ and $\sig_X$ are the mean and variance of $X$, and $\EE[X]$ is the expected value of $X$.  If $\rho_{MB}=1$, then both $\MeB$ and $\mB \overset{\varepsilon}{{\sim}}_F \mM$. Thus, perfect correlation implies supervenience, but supervenience does not imply correlation.

 % Consider the following analogy: one desires to understand the relationship between height and weight.  In such a scenario, height, $h$, could be assumed to be an element from the set of all possible heights, $\mc{H}=(0,\infty)$ cm; and weight, $w$, could be assumed to be an element from the set of all possible weights, $\mc{W}=(0,\infty)$ kgs.   Then, it may be assumed that an agent's height and weight is sampled identically and independently from some \emph{model}, that is $(h,w)\overset{iid}{\sim}\PP[H,W]=F_{HW}$. To gain insight into $F_{HW}$, one might then state hypotheses, collect data, and perform analyses.  Implicit in the exposition of a model is that both objects, height and weight here, are \emph{random objects}.  Note that this does not imply anything in particular about the distribution (which could be a point mass, for instance); rather, it is a formal definition necessary for statistical analysis and empirical investigation.

\subsection{Brain-graphs} % (fold)
\label{sub:brain_graphs}

Given the above notion of statistical supervenience, one can then design a statistical test for supervenience by carefully considering an appropriate model.  For mind-brain supervenience, a model of the brain is paramount, and has historically been a bit of a sticky wicket.  It was not until 1891, that H.~Waldeyer-Hartz first formally proposed the ``neuron doctrine'' \cite{Waldeyer-Hartz1891}, which states that the nervous system is a complex \emph{network} of ``neurons'' (a term he invented), largely based on Ramon y Cajal's work using the Golgi stain \cite{Finger01}.  This doctrine has been central to much of the development of neuroscience and artificial intelligence for over 100 years, including the development of neural network theory \cite{Bishop95} and cognitive science \cite{McClellandRumelhart86}.  Fundamental to nearly all models of the brain (or components therein) since the neuron doctrine's introduction is the existence of neural units, each connected to (some) other neural units.  Therefore, it seems appropriate to assume that a reasonable model of a brain is graph, composed of a set of vertices (or nodes), each representing a neural unit, and a set of edges (arcs) connecting the nodes.  Below, we formalize this concept to enable statistical testing in the graph domain.

% subsection brain_graphs (end)

% section preliminaries (end)

\section{Results}

\subsection{Theoretical results} % (fold)
\label{sub:theoretical_results}

% subsection theoretical_results (end)
If minds statistically supervene on brains, $\MsB$, then two different minds must supervene on two different brains.  This means that there exists a unique mapping from each brain to a single mind.  In other words, one can in principle construct a function $g(b): \mB \mapsto \mM$, that is a deterministic mapping from brains to minds.  It may be the case that subsets of brains form equivalence classes, such that any brain in that subset is mapped to the same mind (see, for example, $b_1$ and $b_2$ in Figure \ref{fig:rel}(A)).  Assuming for the moment that the space of all possible minds is finite, that is $| \mM| < \infty$, then we call any such function a \emph{classifier} (this assumption will later be relaxed).  Let $\mh{m}$ denote the output of a classifier, $g(b)=\mh{m}$.  Define misclassification rate as:
\begin{align}
	L_{\PP}(g) = \PP[g(B) \neq M] = \frac{1}{|\mB| |\mM|}\iint \II\{g(b)\neq m\} db dm
\end{align}
where $\II\{\cdot\}$ indicates the indicator function, taking unity value if its argument is true, and zero otherwise.  $L_{\PP}(g)$ therefore effectively counts the fraction of time $g$ misclassifies $b$. The Bayes optimal classifier $g^*$ minimizes $L_{\PP}(g)$ over all classifiers, that is
\begin{align}
	g^* = \argmin_{g \in \mc{G}} L_{\PP}(g)
\end{align}
where $\mc{G}$ is the set of all possible classifiers.  Thus, the \emph{Bayes error}, or Bayes risk, $L_{\PP}(g^*)$ is the minimum possible misclassification rate. The primary result of casting supervenience is a statistical framework is the following theorem: 
\begin{thm}
\label{thm1} 
$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $L_{\PP}(g^*) = 0$.
\end{thm}
% \begin{thm} 
% 	\label{thm1} 
% 	Statistical supervenience implies that there exists a classifier that is \emph{always correct}, that is, $g(b)=m$, for all $(m,b) \in \mc{M} \times \mc{B}$.  
% \end{thm}

\noindent If minds supervene on brains, then, by the definition of supervenience, there exists a function that maps each brain deterministically to a particular mind.  This means that one could draw a decision boundary between all equivalence classes of brains, each class corresponding to a different mind, and no mind will reside within two different equivalence classes.  Thus, the optimal classifier would correctly find these decision boundaries, and therefore have no opportunity to err. $\square$

This relationship between statistical supervenience and Bayes error can therefore be described concisely: \mbox{$\MsB \Leftrightarrow L_{\PP}(g^*)=0$}.  Thus, the above arguments shows (for the first time to our knowledge) that statistical supervenience and zero Bayes error are equivalent. Further, statistical supervenience can be thought of as a constraint on the possible models.  Specifically, let $\mc{P}_s \subseteq \mc{P}$ be subset of models for which supervenience holds.  Then, $\mc{P}_s = \{\PP[M,B] : L_{\PP}(g^*)=0\} \subseteq \mc{P}$.

\subsection{Hypothesis testing} % (fold)
\label{sub:hypothesis_testing}
